Evaluation of dynamic resource allocation strategies for overlay Apache Spark clusters on HPC \tristan{``Overlay'' and ``Apache Spark'' are redundant}

### 2019-05-08 paper.tex

    Two different job templates were used to implement the two main
    conditions compared in our experiments: the batch submission
    template and the pilot submission template. The batch submission template was
    inspired by the template provided by Compute Canada to launch Spark applications 
    on Slurm \href{https://docs.computecanada.ca/wiki/Apache\_Spark/en}{available here}.
    The template operates as follows: Certain resource requirements are requested by
    the user (e.g. walltime, amount of memory per node, number of CPUs per task, number 
    of nodes and number of tasks per node). Once these resources are allocated, a 
    master is started on one of the requested node resources. Then, after the master
    has successfully started, the workers are started on all nodes. Multiple worker
    instances are started on a single node by setting the \textit{SPARK\_WORKER\_INSTANCES}
    environment variable to the number of tasks per node. Each worker is given as many
     cores as specified by the user in the Slurm resource allocation request.
    After both the masters and the workers have successfully started, the driver is finally
    started. The amount of memory given to each executor corresponds to 95\% of the 
    available memory on the node, to allow for offheap space. 
    The Spark deploy-mode selected for the batch template is client-mode.

    The pilot submission template is similar to that of the batch template, however,
    each pilot will start its own Spark master and worker. Only one pilot,
    the first pilot to attempt to launch the driver, will do so. The reason for 
    which each pilot starts its own master is to ensure the fault tolerance of the 
    masters. In this sort of configuration, should the active master be killed, one
    of the stand-by masters will takeover and the application will be able to 
    resume. Such a configuration is particularly favourable in pilot scheduling 
    scenarios as node failures may be more frequent due to walltime expiration.
    Additionally, the Spark deploy mode of the driver was selected to be cluster deploy mode.
    This would not only allow the driver to be executed directly on one of the workers,
    but also allow us to make the driver fault-tolerant through the
    \textit{supervise} mode, which is only available in cluster deploy
    \TG{A brief section on Spark to explain these modes and features would
    be useful}. As with the masters, it is particularly important to have a
    fault-tolerant driver in pilot-scheduling scenarios due to possible
    walltime expiration \TG{It should be explained before that one of the interests of pilots
    is to ``play'' with shorter walltimes}. \TG{It would also be nice to have runs where pilots 
    have shorter walltimes, to see if queuing times are shorter too.}
    Should pilots be idle for a certain duration, the
    pilots will terminate themselves such as to not hog resources.

@@

    \subsection{Added value of pilot scheduling}
        To determine if there are any performance benefits of using pilot over 
        batch scheduling, we needed to compare both strategies given various resource
        requirements. It is expected that large batch requests will stay in the 
        resource queue longer than multiple pilot requests that ultimately use up
        the same amount of resources as each individual pilot requests less resources at
        a time. We therefore used four resource configurations to investigate this
        hypothesis (Table~\ref{table:dedicatednodes}).

        \begin{table}                                                                    
\centering                                                                       
\begin{tabular}{c|c|c}                                                             
\rowcolor{headcolor}                                                             
Configuration & RAM (GB) & Tasks \\                               
\hline                                                                           
1 & 112 & 16 \\                                               
2 & 224 & 32 \\                                               
3 & 336 & 48 \\
4 & 448 & 64 \\
\end{tabular}                                                                    
\setlength{\belowcaptionskip}{-10pt}                                             
\caption{Resource configurations}                                                    
\label{table:dedicatednodes}                                                            
\end{table} 
        \todo{execution of a batch cluster vs dynamic (pilot) cluster}
    \subsection{Robust masters}
        \todo{scala standalone vs our pyspark workaround. kill masters in experiments}
    \subsection{Checkpointing}
        \todo{metric for determining how often to checkpoint based on cluster size}
    \subsection{Job arrays}
        \todo{need to kill idle workers. may not want all workers to be running at once.}
    \subsection{Example application}
        \todo{incrementation with varying task durations}
\section{Results}\label{sec:Results}
\section{Discussion}\label{sec:discussion}

>>>>

    Two different job templates were used to implement the two main
    conditions compared in our experiments: the batch submission
    template and the pilot submission template. The batch submission template was
    inspired by the template provided by Compute Canada to launch Spark applications 
    on Slurm \href{https://docs.computecanada.ca/wiki/Apache\_Spark/en}{available here}.
    The template operates as follows: Certain resource requirements are requested by
    the user (e.g. walltime, amount of memory per node, number of CPUs per task, number 
    of nodes and number of tasks per node). Once these resources are allocated, a 
    master is started on one of the requested node resources. Then, after the master
    has successfully started, the workers are started on all nodes. Multiple worker
    instances are started on a single node by setting the \texttt{SPARK\_WORKER\_INSTANCES}
    environment variable to the number of tasks per node. Each worker is given as many
     cores as specified by the user in the Slurm resource allocation request.
    After both the masters and the workers have successfully started, the driver is finally
    started. The amount of memory given to each executor corresponds to 95\% of the 
    available memory on the node, to allow for offheap space. 
    The Spark deploy-mode selected for the batch template is client-mode.

    The pilot submission template is similar to that of the batch template, however,
    each pilot will start its own Spark master and worker. Only one pilot,
    the first pilot to attempt to launch the driver, will do so. The reason for 
    which each pilot starts its own master is to ensure the fault tolerance of the 
    masters. In this sort of configuration, should the active master be killed, one
    of the stand-by masters will takeover and the application will be able to 
    resume. Such a configuration is particularly favourable in pilot scheduling 
    scenarios as node failures may be more frequent due to walltime expiration.
    Additionally, the Spark deploy mode of the driver was selected to be cluster deploy mode.
    This would not only allow the driver to be executed directly on one of the workers,
    but also allow us to make the driver fault-tolerant through the
    \texttt{supervise} mode, which is only available in cluster deploy
    \TG{A brief section on Spark to explain these modes and features would
    be useful}. As with the masters, it is particularly important to have a
    fault-tolerant driver in pilot-scheduling scenarios due to possible
    walltime expiration \TG{It should be explained before that one of the interests of pilots
    is to ``play'' with shorter walltimes}. \TG{It would also be nice to have runs where pilots 
    have shorter walltimes, to see if queuing times are shorter too.}
    Should pilots be idle for a certain duration, the
    pilots will terminate themselves such as to not hog resources.

@@

    \subsection{Added value of pilot scheduling}
        \begin{table}                                                                    
            \centering                                                                       
            \begin{tabular}{c|c|c}                                                             
            \rowcolor{headcolor}                                                             
            Configuration & RAM (GB) & Tasks \\                               
            \hline                                                                           
            1 & 112 & 16 \\                                               
            2 & 224 & 32 \\                                               
            3 & 336 & 48 \\
            4 & 448 & 64 \\
            \end{tabular}                                                                    
            \setlength{\belowcaptionskip}{-10pt}                                             
            \caption{Resource configurations}                                                    
            \label{table:dedicatednodes}                                                            
        \end{table}

        \begin{table*}                                                                   
        \centering                                                                       
        \begin{tabular}{c|cccccc}                                                   
          \rowcolor{headcolor}                                                           
          \multicolumn{7}{c}{Configuration 1}\\                      
          \hline                                                                         
          \rowcolor{headcolor}                                                           
          execution mode & nodes/job & RAM (G) & CPUs per task & tasks/node & walltime & task delay (s) \\                             
          \hline
          Batch & 1 & 112 & 1 & 16 & 2h30 & 45 \\
          8 pilots & 1 & 14 & 1 & 2 & 2h30 & 45 \\
          16 pilots & 1 & 7 & 1 & 1 & 2h30 & 45 \\

          \hline                                                                           
          \multicolumn{7}{c}{}\\

          \rowcolor{headcolor}                                                           
          \multicolumn{7}{c}{Configuration 2}\\                      
          \hline                                                                         
          \rowcolor{headcolor}                                                           
          execution mode & nodes/job & RAM (G) & CPUs per task & tasks/node & walltime & task delay (s) \\                             
          \hline
          Batch & 2 & 112 & 1 & 16 & 2h30 & 90 \\
          8 pilots & 1 & 28 & 1 & 4 & 2h30 & 90 \\
          16 pilots & 1 & 14 & 1 & 2 & 2h30 & 90 \\

          \hline                                                                           
          \multicolumn{7}{c}{}\\

          \rowcolor{headcolor}                                                           
          \multicolumn{7}{c}{Configuration 3}\\                      
          \hline                                                                         
          \rowcolor{headcolor}                                                           
          execution mode & nodes/job & RAM (G) & CPUs per task & tasks/node & walltime & task delay (s) \\                             
          \hline
          Batch & 3 & 112 & 1 & 16 & 2h30 & 120 \\
          8 pilots & 1 & 42 & 1 & 6 & 2h30 & 120 \\
          16 pilots & 1 & 21 & 1 & 3 & 2h30 & 120 \\

          \hline                                                                           
          \multicolumn{7}{c}{}\\

          \rowcolor{headcolor}                                                           
          \multicolumn{7}{c}{Configuration 4}\\                      
          \hline                                                                         
          \rowcolor{headcolor}                                                           
          execution mode & nodes/job & RAM (G) & CPUs per task & tasks/node & walltime & task delay (s) \\                             
          \hline
          Batch & 4 & 112 & 1 & 16 & 2h30 & 180 \\
          8 pilots & 1 & 56 & 1 & 8 & 2h30 & 180 \\
          16 pilots & 1 & 28 & 1 & 4 & 2h30 & 180 \\

          \hline                                                                           
          \multicolumn{7}{c}{}\\                                                        
        \end{tabular}                                                                    
        \setlength{\belowcaptionskip}{-10pt}                                             
        \caption{Experimental conditions}
        \label{table:conditions}                                                        
        \end{table*}                                                                                   
        To determine if there are any performance benefits to using pilot over 
        batch scheduling, we needed to compare both strategies given various resource
        requirements. It is expected that large batch requests will stay in the 
        resource queue longer than multiple pilot requests that ultimately use up
        the same amount of resources as each individual pilot requests less resources at
        a time. We therefore used four resource configurations to investigate this
        hypothesis (Table~\ref{table:dedicatednodes}).

       For batch, we requested 1 to 4 dedicated nodes, depending on resource 
       configuration, each with 112~G of RAM, 1 core per task and 16 tasks per 
       node. On the other hand, for all configurations, we ran our experiments
       with 8 and 16 pilots. Experimental conditions can be seen in Table~\ref{table:conditions}.

       As the focus of our experiments here is to measure the impact pilot scheduling
       has on queueing times, we wanted to ensure that our walltime estimates would
       be consistent for each experimental condition. Therefore, we adjusted task 
       delay based on the maximum level of parallelism to ensure that walltimes would
       not need to be readjusted for each configuration. Given utmost parallelism,
       the sleep delay added would amount to 1 hour of processing time regardless
       of configuration. The real processing time of the application, however, 
       would differ between configurations as we did not account for task duration
       in the sleep delay. However, task duration was not expected to signicantly impact
       overall processing time, but nevertheless, we set application walltime to 
       2 hours and 30 minutes for all configurations to account for additional
       processing time that may be incurred in applications with lower levels of 
       parallelism.

       For each configuration, all three execution modes were executed in parallel
       to ensure that the status of the overall cluster was the same when the 
       different execution modes were launched. Furthermore, the order of the execution modes
       for each configuration was randomized to ensure that SLURM job request
       order could not have affected results.

       There were 15 repetitions in total for each configuration, and the order
       in which the configurations were launched was also randomized. This was to 
       account for any system variability that can occur, particularly in production
       HPC clusters.

       Different clusters may have different resource configurations, number of 
       users and scheduling policies. Therefore, we executed all 15 repetitions on
       both Cedar and B\'eluga to determine how much our results differ between
       two distinct clusters.

    \subsection{Pipeline robustness}
        While nodes failures can occur for various reasons, underestimation of 
        walltime may lead to early node termination in pilot scheduling systems.
        Therefore, it is crucial for the overlay cluster within the pilot system
        to be robust to node failures. In this series of experiments, we investigate
        Spark's fault-tolerance with respect to worker, master and driver termination.

        In our configuration, the driver is running in cluster deploy-mode, resulting
        in it being executed directly on a worker. This feature alone should not make
        it fault-tolerant to failures. However, using this feature coupled with the 
        \texttt{--supervise} flag, will.

        \todo{scala standalone vs our pyspark workaround. kill masters in experiments}
    \subsection{Checkpointing}
        \todo{metric for determining how often to checkpoint based on cluster size}
    \subsection{Job arrays}
        \todo{need to kill idle workers. may not want all workers to be running at once.}
    \subsection{Example application}
        \todo{incrementation with varying task durations}
\section{Results}\label{sec:Results}
\section{Discussion}\label{sec:discussion}
